\chapter{Conclusion}
\label{cha:conclusion}

In this report the Kerr effect and wavemixing have been investigated. 

The Kerr effect on the refractive index was worked out from the second nonlinear term of the
polarization. From this the self-focusing effect could be explained by the in-homogeneous
refractive index composition.  And the critical intensity for this effect to conquer diffractive
spreading was presented for the cylindrical and the Gaussian beam, which was basis wave for the
numerical work.

Here the point of departure was the dimensionless waveequation, where the second order derivatives
was approximated by the central finite difference. Then using the Crank-Nicolson method the
problem was reformulated to a five diagonal matrix which should be solved for each time
step. The implementation of the Crank-Nicolson was numerically unstable. Though a tendency of
centered amplification was observed, either energy conservation broke down, or strong oscillations
manifested at the center.


Since there was issues with the numerical solutions to the partial differential equations a
qualitative explanation of the three wave mixing was presented. The second order term couples the
two waves together, which then couple back to the electric field creating a third wave. The
frequency will be the sum or difference of the initial frequencies, assuming that the phase matching
constraint are fulfilled. Finally examples of the up-conversion in praxis with efficiency's up to
$93.7\%$ are presented.


